---
title: "Parse all the KGML files for human KEGG pathways and save as csv files"
date: "`r Sys.time()`"
output: html_document
---

Load libraries:

```{r}
library(KEGGgraph) ##use this to convert KGML to data frame
library(KEGGREST) ##use this to convert KEGG gene IDs to NCBI gene IDs - no longer needed
library(org.Hs.eg.db) ##use this to convert gene IDs to gene names
library(dplyr)
library(igraph) ##use this to convert to graph object
```

Loop through all KGML files representing human KEGG pathways, convert them into data frames and save them as csv files in ../All_human_pathways_csv. Also save the network diagrams in ../All_human_pathways_figs (this is just for checking). I also converted the KEGG gene IDs to regular gene IDs. Note that some pathways did not convert properly for various reasons and thus were not saved as csv/png files (could be there were no genes or proteins or there were unusual schemas).

```{r}
##get all the KGML xml files in ../Pathways_KGML and convert them to data frames
all_kgml <- list.files("../All_human_pathways_KGML")
for(file_kgml in all_kgml)
{
  file_path_kgml <- paste("../All_human_pathways_KGML",file_kgml, sep="/")
  
  ##convert KGML to data frame
  current_df <- parseKGML2DataFrame(file_path_kgml)
  
  if(nrow(current_df) > 0)
  {
    current_df$from <- as.character(current_df$from)
    current_df$to <- as.character(current_df$to)
    ##get all KEGG IDs for nodes
    all_nodes <- unique(c(current_df[,1], current_df[,2]))
    
    ##first get rid of the "hsa:" part in front of the IDs
    all_nodes_gid <- gsub("hsa:","",all_nodes)
    names(all_nodes_gid) <- all_nodes

    if(sum(!is.na(all_nodes_gid)) > 0)
    {
      length(unique(all_nodes))
      length(unique(all_nodes_gid))
      head(all_nodes)
      head(all_nodes_gid)
      
      gid_to_name <- as.character(org.Hs.egSYMBOL)
      all_nodes_gene_symbol <- gid_to_name[all_nodes_gid] 
      head(all_nodes_gene_symbol)
      length(all_nodes_gene_symbol)

      all_nodes_gene_symbol <- gid_to_name[all_nodes_gid] 
      
      ##now put in gene names in the data frame
      ##only keep KEGG IDs that do get mapped 
      ##(the other IDs are probaly things like pathways _not_ genes/proteins)
      current_df <- dplyr::filter(current_df, 
                                  from %in% names(all_nodes_gid),
                                  to %in% names(all_nodes_gid))
      
      current_df$from <- all_nodes_gene_symbol[all_nodes_gid[current_df$from]]
      current_df$to <- all_nodes_gene_symbol[all_nodes_gid[current_df$to]]
      head(current_df)
      
      graph <- graph_from_data_frame(current_df)
      V(graph)$size <- 8
      V(graph)$label.cex <- 0.6
      
      png_name <- paste("../All_human_pathways_figs",gsub("(.xml)|(.kgml)", ".png",file_kgml),sep="/")
      
      png(filename=png_name)
      try(plot(graph,
               edge.arrow.size=0,
               layout=layout_with_kk))
      dev.off()
      
      csv_name <- paste("../All_human_pathways_csv",gsub("(.xml)|(.kgml)", ".csv",file_kgml),sep="/")
      
      write.csv(x=current_df, file=csv_name)
      
    }  
  }
}
```

Also obtain all the connections between genes/proteins based on the KEGG pathways and create a data frame with them, saved in ../objs/connections_KEGG.RData. Furthermore, create a list of dataframes, saved in the same place.
```{r}
##load the KEGG pathways and store them in a single data frame
all_KEGG <- list.files("../All_human_pathways_csv")
connections_KEGG <- data.frame()
list_paths_KEGG <- list()
for(KEGG_path in all_KEGG)
{
  KEGG_path_df <- read.csv(paste("../All_human_pathways_csv",
                                 KEGG_path, sep="/"))
  connections_KEGG <- rbind(connections_KEGG,
                            KEGG_path_df)  
  KEGG_path_name <- gsub(".csv","",KEGG_path)
  list_paths_KEGG[[KEGG_path_name]] <- KEGG_path_df[,-1]
}
##get rid of first column
connections_KEGG <- connections_KEGG[,-1]

save(list="connections_KEGG", file="../objs/connections_KEGG.RData")
save(list="list_paths_KEGG", file="../objs/list_paths_KEGG.RData")

```
