---
title: "Therapy recommendations using biological networks: A breast cancer example"
output: html_document
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(shiny.sanitize.errors = FALSE)
```

<span style="color:red">
<strong>Warning!</strong> The following tool is for research purposes only. It is not intended for clinical care. </span>

### Load libraries:
```{r include=TRUE, message=FALSE}
library(utils)
library(shiny)
library(shinyBS)
library(ggvis)
library(networkD3)
library(igraph)
library(kableExtra)
library(knitr)
library(dplyr)
library(rDGIdb)
library(networkViz) ##this package is available at https://github.com/jkanche/nfpmShinyComponent
```

### Load objects:
```{r}
load("../data/example_CDGnet_database_inputs.RData")

ls()
```
The example_CDGnet_database_inputs.RData file only has the drugs_PO_FDA_biomarkers, drugs_PO_FDA_targets, FDA_approved_drugs objects. To obtain the DrugBank_targets,
KEGG_oncogenes, and list_paths_KEGG objects, see the files in the code_process_KEGG_DrugBank directory.

### Source functions:
```{r}
source("../code/functions.R")
```

### Read in the example data:

This is example data for a putative patient with ER+ breast cancer (ESR1 overexpression), who also has FGFR1 overexpression:
```{r}
MP <- read.table("../data/example_input.tsv",
                 sep="\t",
                 header = TRUE)

MP
```

### Category 1 recommendations:

Category 1 therapy recommendations consist of FDA-approved drugs _for these alterations in this cancer type_, i.e. the alterations represent biomarkers for this cancer type:
```{r}
Type1 <- get_cat_1_2(MP,
                     cancer_type = "Breast cancer",
                     drugs_PO_FDA_biomarkers)
Type1
```

### Category 2 recommendations:

Category 2 therapy recommendations consist of FDA-approved drugs _for these alterations in other cancer types_, i.e. the alterations represent biomarkers for other cancer types:
```{r}
Type2 <- get_cat_1_2(MP,
                     cancer_type = "Breast cancer",
                     drugs_PO_FDA_biomarkers,
                     cat2 = "yes")
Type2
```

### Category 3 recommendations:

Category 3 therapy recommendations consist of drugs which have as targets the input genes/proteins or as biomarkers/targets other genes/proteins that are _downstream of the input oncogenes_ when considering the _pathway corresponding to this tumor type_. We only consider
the option of FDA-approved targeted cancer therapies. First get the recommendations:
```{r}
Type3 <- get_cat_3_4(MP, ##input data frame
                     cancer_type = "Breast cancer", ##character string
                     cat_drugs = "Only FDA-approved targeted therapies for cancer", ##character string - whether only targeted cancer therapies, all FDA approved therapies, or all drugs in DrugBank
                     list_paths = list_paths_KEGG, ##list of pathways
                     KEGG_oncogenes, ##data frame with oncogenes and cancer type
                     drugs_PO_FDA_biomarkers, ##FDA-approved precision oncology drugs with listed biomarkers
                     drugs_PO_FDA_targets, ##FDA-approved precision oncology drugs with targets
                     drug_targets = DrugBank_targets, ##data frame of targets for specific genes/proteins - right now, have DrugBank_targets
                     FDA_approved_drugs, ##list of FDA-approved drugs)
                     Type1 = Type1, ##results of get_cat_1_2
                     Type2 = Type2, ##results of get_cat_1_2
                     Type3 = NULL, ##results of get_cat_3_4 (only use in cat4 == "yes)
                     cat4 = "no",
                     subtype = "activation")
dim(Type3$drugs_mat)
head(Type3$drugs_mat)
```

Now visualize the network (must use console for this):
```{r}
res <- dataParser(MP, Type1, Type2, Type3$drugs_mat)

chart <- networkViz:::NfpmViz(data = res)
chart
```

### Category 4 recommendations:

Category 4 therapy recommendations consist of drugs which have as targets the input genes/proteins or as biomarkers/targets other genes/proteins that are _downstream of the input oncogenes_ when considering the _pathway corresponding to other tumor types_. We only consider
the option of FDA-approved targeted cancer therapies:
```{r}
Type4 <- get_cat_3_4(MP, ##input data frame
                     cancer_type = "Breast cancer", ##character string
                     cat_drugs = "Only FDA-approved targeted therapies for cancer", ##character string - whether only targeted cancer therapies, all FDA approved therapies, or all drugs in DrugBank
                     list_paths = list_paths_KEGG, ##list of pathways
                     KEGG_oncogenes, ##data frame with oncogenes and cancer type
                     drugs_PO_FDA_biomarkers, ##FDA-approved precision oncology drugs with listed biomarkers
                     drugs_PO_FDA_targets, ##FDA-approved precision oncology drugs with targets
                     drug_targets = DrugBank_targets, ##data frame of targets for specific genes/proteins - right now, have DrugBank_targets
                     FDA_approved_drugs, ##list of FDA-approved drugs)
                     Type1 = Type1, ##results of get_cat_1_2
                     Type2 = Type2, ##results of get_cat_1_2
                     Type3 = Type3$drugs_mat, ##results of get_cat_3_4 (only use in cat4 == "yes)
                     cat4 = "yes",
                     subtype = "activation")
dim(Type4$drugs_mat)
head(Type4$drugs_mat)
```

Now visualize the network (must use console for this):
```{r}
res <- dataParser(MP, Type1, Type2, Type4$drugs_mat)

chart <- networkViz:::NfpmViz(data = res)

chart$render_component(shiny=FALSE)
```

Session info:
```{r}
sessionInfo()
```
